---
title: 'Bayesian Hierarchial Modeling of Historical Irrigaiton Patterns'
author: 'Rachel Ledig'
date: '`r ifelse(knitr::is_html_output(), "Supervisors: Dr. Tobias Krueger & Dr. Christian Schleyer <br> July 2021", "July 2021")`'
institution: 'University of Humboldt'
division: 'Albrecht Daniel Thaer Institute for Agricultural and Horticultural Sciences'
advisor: 'Dr. Tobias Krueger'
altadvisor: 'Dr. Christian Schleyer'
department: 'Faculty of Life Sciences'
supervisor: 'Dr. Fabian Stenzel'
degree: 'Master of Science'
matriculation: '602220'
email: 'ledigrac@hu-berlin.de'
bday: 'Nov 17, 1991'
bplace: 'California, USA'
knit: bookdown::render_book
site: bookdown::bookdown_site
header-includes: 
   - \usepackage{wrapfig}
always_allow_html: true

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True
  
# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.

output:
# COMMENT OUT THE GIFS PLOTLY BLAH BLAHHHH
#  thesisdown::thesis_pdf: default 
  thesisdown::thesis_gitbook:     
    css: 
    anchor_sections: false
    config: 
      fontsettings:
        theme: 
        family: 
        size:
      sharing:
        facebook: no
        github: yes
        twitter: no
        linkedin: no
        all: ['twitter', 'linkedin', 'facebook']
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abbreviations: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abbreviations.Rmd")), collapse = "\n  ")`'
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-acknowledgements.Rmd")), collapse = "\n  ")`'
dedication: |
  To Mom, Dad, Sis, and Fi.
  
# Specify the location of the bibliography below
bibliography: references.bib
# Download your specific csl file and refer to it in the line below.
link-citations: yes
colorlinks: no  
csl: csl/apa.csl
lot: true
lof: true
---

```{=html}
<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->
```
```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

```{=html}
<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->
```
```{r eval= !knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

```{=html}
<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->
```
# Introduction {#intro}

The practice of irrigation, and its subsequent impacts, have far reaching effects for the global economy, food and water supply, and environment. Irrigated land accounts for 20% of the global cultivated land area which derives 40% of the global food supply [@neumannExploringGlobalIrrigation2011, @portmannMIRCA2000GlobalMonthly2010a], 70% of water withdrawls from ground water aquifers are used for irrigation [@faoStateWorldLand2011], and

## Objectives {#objectives}

Ultimately, the agricultural system will continue to demand more water while competing with other sectors over decreasing water availability in the coming years. With this pressure, farmers will aim to be as efficient as possible by increasing their yields which, in certain areas and under certain conditions, can be done by irrigating their fields. Understanding where, when, and why farmers decide to begin irrigating is important to understand water demands now and in the future. Unfortunately, irrigation expansion is seldom studied, with few studies using statistical modeling techniques to investigate the role that biophysical, economic, and political factors play in the spread of irrigation infrastructure in the present. None do so historically. This leads to the first and overarching research question:

<center>

*Which factors (drivers) have influenced the expansion of irrigation in both time and space over the past 100 years?*

</center>

By understanding under which conditions irrigation infrastructure is built could be coupled with projections of future climate and economic scenarios, allowing us to get a better idea of scales of agricultural production, land use, and water needs in the coming years. In order to eventually extrapolate to the future, predictive accuracy needs to be addressed. This can be done by separating data into training and testing datasets, model tuning, and error estimation.

Several other questions smaller questions arise in regard to drivers of expansion which are interesting to investigate. Such as:

-   Which has a larger influence on irrigation expansion patterns, biophysical or socio-economic factors?

-   Do rates of irrigation expansion differ depending on crop type?

Both of these questions have interesting policy implications, should large trends be present in the model and analysis.

The time series component of this data set also gives rise to some interesting questions.

-   Expansion rates of total irrigated area differ temporally, what could be the drivers for this difference?

-   Is there a temporal delay in irrigation expansion based on the influence of certain drivers, and if so how much?
